Reproducible Research - Peer Assignment # 1
========================================================

## Loading and preprocessing the data
- Load activity data and transform the date variable to date class

```{r}
library(stats)
setwd("~/Dropbox/Code")
data = read.csv("./rrProject1/activity.csv", header = TRUE)
data$date = as.Date(data$date, "%Y-%m-%d")
```

## What is mean total number of steps taken per day?
- Make a histogram of the total number of steps taken each day

```{r fig.width=7, fig.height=6}
stepsbyday = aggregate(steps ~ date, data = data, sum)
hist(stepsbyday$steps)
```

- Calculate and report the mean and median total number of steps taken per day
```{r}
meanSteps = mean(stepsbyday$steps, na.rm = TRUE) 
medianSteps = median(stepsbyday$steps, na.rm = TRUE)
```
The mean number of steps taken per day is `r meanSteps` and the median number of steps
take per day is `r medianSteps`.

## What is the average daily activity pattern?

- Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r}
library(ggplot2)
avgSteps  = data.frame(xtabs(steps~interval,aggregate(steps~interval,data,mean)))
qplot(interval, Freq, data = avgSteps, ylab = "Average Steps",
      xlab = "5-Minute Daily Interval") + labs(title = "Average Number of Steps Taken by 5-Minute Interval")
```

- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
``` {r}
maxInterval = which.max(avgSteps$Freq)
maxSteps = max(avgSteps$Freq)
```

Interval `r maxInterval` contains the maximum number of steps on average of `r maxSteps`.

## Imputing missing values

- Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

``` {r}
totalRecords = length(data$steps)
completeRecords = length(na.omit(data$steps))
missingRecords = totalRecords - completeRecords
```

Of the `r totalRecords` total records in the dataset, only `r completeRecords` contain
valid data. The difference, `r missingRecords`, is the total number of missing values.

- Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

Replace missing step values with the mean for a given 5-minute interval across the entire observation period

``` {r}
data$steps2 = data$steps
for (i in 1:length(data$steps))
        if(is.na(data$steps[i])) {
                data$steps2[i] = mean(data$steps, na.rm = TRUE)
        }
```

- Create a new dataset that is equal to the original dataset but with the missing data filled in.

``` {r}
dataNew = data.frame(steps = data$steps2, date = data$date, interval = data$interval)
```

- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
``` {r}
stepsbydayNew = aggregate(steps ~ date, data = dataNew, sum)
hist(stepsbydayNew$steps)
mean(stepsbydayNew$steps, na.rm = TRUE) 
median(stepsbydayNew$steps, na.rm = TRUE)
```

## Are there differences in activity patterns between weekdays and weekends?

- For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

``` {r}
day = weekdays(dataNew$date)
dayType = vector()
for (item in day) { 
        if (item == "Saturday" || item == "Sunday") {
        dayType = append(dayType, "weekend")
        } else {
        dayType = append(dayType, "weekday")        
        }   
}
dataNew$dayType = factor(dayType) 

```
- Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:

``` {r}
avgStepsNew  = data.frame(xtabs(steps~interval+dayType,aggregate(steps~interval+dayType,dataNew,mean)))
qplot(interval, Freq, data = avgStepsNew, facets = dayType~.)

```